---
title: "Adding custom adverse event filters"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Adding custom adverse event filters}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  eval = FALSE,
  collapse = TRUE,
  comment = "#>"
)
```


{dv.clinlines} offers optional local filters for adverse event data which will be added to the bottom of the sidebar. The following filters are available:

-   `serious_AE`: Offers the possibility to filter adverse events to show only serious/non-serious adverse events.
-   `soc`: Offers the possibility to filter adverse events by System Organ Class.
-   `pref_term`: Offers the possibility to filter adverse events by Preferred Term.
-   `drug_rel_AE`: Offers the possibility to filter adverse events to show only drug related/non drug related adverse events.

Activate local filters by adding their names through a list to the `filter` parameter of the `mod_clinical_timelines()` call. Use `NULL` (the default) to turn all filters down. To specify one or multiple local adverse event filters, you need to provide a list with the following elements:

-   `data_name`: Character name of the adverse events dataset of your loaded data list. Must be a single value. (Mandatory.)
-   `label`: Character value which is exactly the same as the name for the adverse events event defined in the `mapping` parameter of `mod_clinical_timelines()`.
-   `serious_AE`: Character name of the variable that holds Y/N flags for serious adverse events. Must be a single value.
-   `soc`: Character name of the variable that holds system organ classes. Must be a single value.
-   `pref_term`: Character name of the variable that holds preferred terms. Must be a single value.
-    `drug_rel_AE`: Character name of the variable that holds Y/N flags for causality. Must be a single value.

Wrap this list with another list. The inner list must be named `ae_filter`. The outer list should be assigned to the `filter` parameter of `mod_clinical_timelines()`. 


Find an example of the code structure below:
```{r}
# To deactivate local filters
filter <- NULL

# To activate all local adverse event filters
filter <- list(
  ae_filter = list(
    data_name = "adae", # mandatory - as defined in the data list
    label = "Adverse Events", # mandatory - as defined in the mapping parameter
    soc = "AESOC",
    serious_AE = "AESER",
    pref_term = "AEDECOD",
    drug_rel_AE = "AEREL"
  )
)
```


For multiple filters, filtering will be performed using AND condition.
<br>
<br>
The code below shows a module list containing a Clinical Timelines module with only three local adverse event filters:

```{r ae_filter}
module_list <- list(
  "Clinical Timelines" = mod_clinical_timelines(
    module_id = "mod1",
    basic_info = list(
      data = "adsl",
      trt_start = "TRTSDT",
      trt_end = "TRTEDT",
      icf_date = "RFICDT"
    ),
    mapping = list(
      adsl = list(
        "Treatment Start" = list(
          start_dt_var = "TRTSDT",
          end_dt_var = NULL,
          start_dy_var = NULL,
          end_dy_var = NULL,
          detail_var = NULL
        ),
        "Treatment End" = list(
          start_dt_var = "TRTEDT",
          end_dt_var = NULL,
          start_dy_var = NULL,
          end_dy_var = NULL,
          detail_var = NULL
        )
      ),
      adae = list(
        "Adverse Events" = list(
          start_dt_var = "AESTDTC",
          end_dt_var = "AEENDTC",
          start_dy_var = NULL,
          end_dy_var = NULL,
          detail_var = "AEDECOD"
        )
      )
    ),
    drug_admin = list(
      name = "exp",
      start_var = "EXSTDTC",
      end_var = "EXENDTC",
      detail_var = "EXTRT",
      label = "Drug Administration",
      exp_dose = "EXDOSE",
      exp_dose_unit = "EXDOSU"
    ),
    filter = list(
      ae_filter = list(
        data_name = "adae",
        label = "Adverse Events",
        soc = "AESOC",
        serious_AE = "AESER",
        drug_rel_AE = "AEREL"
        # Note: The pref_term filter is not specified in this example.
      )
    )
  )
)
```
